---
title: "Data Exploration"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
always_allow_html: true
---

During our exploratory analysis, we first examined COVID-19 and influenza case and vaccine trends separately to better understand the data we were working with. 

```{r setup, include = FALSE}
library(tidyverse)
library(dplyr)
library(readxl)
library(haven)
library(ggplot2)
library(patchwork)
library(leaflet)


knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

### Data Cleaning

Loading the COVID data from NYC Health.
```{r, message = FALSE}
caserate_by_zip = read_csv("./data/caserate-by-modzcta.csv")
vax_by_boro_age = read_csv("./data/coverage-by-boro-age.csv")
vax_by_boro_demo = read_csv("./data/coverage-by-boro-demo.csv")
data_by_day = read_csv("./data/data-by-day.csv")
doses_by_day = read_csv("./data/doses-by-day.csv")
hosprate_by_zip = read_csv("./data/hosprate-by-modzcta.csv")
testrate_by_zip = read_csv("./data/testrate-by-modzcta.csv")
vax_by_zip_adults = read_csv("./data/coverage-by-modzcta-adults.csv")

boroughs_by_sex = read_csv("./data/boroughs-by-sex.csv")
boroughs_by_age = read_csv("./data/boroughs-by-age.csv")
boroughs_by_race = read_csv("./data/boroughs-by-race.csv")


zip_lat_lon = 
  read.table("./data/us_zipcode_lat_lon.txt", sep = ",", header = TRUE) %>% 
  janitor::clean_names() %>% 
  rename(zipcode = zip) %>% 
  filter(zipcode %in% c(10001:11697))
```
From the datasets above, we will likely compare the following pairs: caserate/hosprate/testrate_by_zip; data/doses_by_day; vax_by_boro_age/demo



Cleaning caserate_by_zip:
```{r}
caserate_by_zipcode = 
  caserate_by_zip %>% 
  janitor::clean_names() %>% 
  select(week_ending, caserate_10001:caserate_10280) %>% 
  pivot_longer(
    caserate_10001:caserate_10280,
    names_to = "zipcode", 
    names_prefix = "caserate_",
    values_to = "case_rate"
  ) %>% 
  mutate(
    week_ending = as.Date(week_ending, "%m/%d/%Y")) 

caserate_by_boro = 
  caserate_by_zip %>% 
  janitor::clean_names() %>% 
  select(week_ending, caserate_city:caserate_si) %>% 
  pivot_longer(
    caserate_city:caserate_si,
    names_to = "boro", 
    names_prefix = "caserate_",
    values_to = "case_rate"
  ) %>% 
  mutate(
    week_ending = as.Date(week_ending, "%m/%d/%Y"),
    boro = recode(boro, "bx" = "Bronx", "bk" = "Brooklyn", "mn" = "Manhattan", "qn" = "Queens", "si" = "Staten Island")) 
```


Cleaning testrate_by_zip:
```{r}
testrate_by_zipcode = 
  testrate_by_zip %>% 
  janitor::clean_names() %>% 
  select(week_ending, testrate_10001:testrate_10280) %>% 
  pivot_longer(
    testrate_10001:testrate_10280,
    names_to = "zipcode", 
    names_prefix = "testrate_",
    values_to = "test_rate"
  ) %>% 
  mutate(
    week_ending = as.Date(week_ending, "%m/%d/%Y")) 

testrate_by_boro = 
  testrate_by_zip %>% 
  janitor::clean_names() %>% 
  select(week_ending, testrate_city:testrate_si) %>% 
  pivot_longer(
    testrate_city:testrate_si,
    names_to = "boro", 
    names_prefix = "testrate_",
    values_to = "test_rate"
  ) %>% 
  mutate(
    week_ending = as.Date(week_ending, "%m/%d/%Y"),
    boro = recode(boro, "bx" = "Bronx", "bk" = "Brooklyn", "mn" = "Manhattan", "qn" = "Queens", "si" = "Staten Island")) 
```


Cleaning hosprate_by_zip:
```{r}
hosprate_by_zipcode = 
  hosprate_by_zip %>% 
  janitor::clean_names() %>% 
  select(date, hosprate_10001:hosprate_10280) %>% 
  pivot_longer(
    hosprate_10001:hosprate_10280,
    names_to = "zipcode", 
    names_prefix = "hosprate_",
    values_to = "hosp_rate"
  )

hosprate_by_boro = 
  hosprate_by_zip %>% 
  janitor::clean_names() %>% 
  select(date, hosprate_bronx:hosprate_citywide) %>% 
  pivot_longer(
    hosprate_bronx:hosprate_citywide,
    names_to = "boro", 
    names_prefix = "hosprate_",
    values_to = "hosp_rate"
  ) %>% 
  mutate(
    boro = recode(boro, "bronx" = "Bronx", "brooklyn" = "Brooklyn", "manhattan" = "Manhattan", "queens" = "Queens", "staten_island" = "Staten Island")
  )
```


Cleaning vax_by_boro_age/demo:
```{r}
vax_by_boro_age_df = 
  vax_by_boro_age %>% 
  janitor::clean_names() %>% 
  filter(age_group %in% c("'18-24", "'25-34", "'35-44", "'45-54", "'55-64", "'65-74", "'75-84", "'85+") )

vax_by_boro_demo_df = 
  vax_by_boro_demo %>% 
  janitor::clean_names() %>% 
  filter(age_group %in% c("'18-44", "'45-64", "'65+", "All ages") )
```

Cleaning data/doses_by_day:
```{r}
doses_by_day_df = 
  doses_by_day %>% 
  janitor::clean_names()

data_by_day_df = 
  data_by_day %>% 
  janitor::clean_names() %>% 
  mutate(
    date_of_interest = as.Date(date_of_interest, "%m/%d/%Y")
  )
```

Cleaning vax_by_zip_adults:
```{r}
vax_by_zip_adults_df = 
  vax_by_zip_adults %>% 
  janitor::clean_names() %>% 
  rename(boro = borough, zipcode = modzcta)
```

## Exploratory Analysis: COVID-19 Rates & Vaccinations
```{r, echo = FALSE, message = FALSE, warning = FALSE}
caserate_by_boro_plot = 
  caserate_by_boro %>% 
  filter(boro != "city") %>% 
  ggplot(aes(x = week_ending, y = case_rate, color = boro)) +
  geom_line() +
  labs(
    title = "Weekly Case Rate by Borough",
    x = "",
    y = "Case Rate"
  )

caserate_by_boro_plot
```


This plot shows the weekly case rate by NYC borough, beginning the week starting on August 9, 2020. The plot shows a similar trend in weekly rates across all 5 boroughs. Cases peaked for all boroughs from January through February of 2021. Boroughs experienced a smaller peak in cases in April 2021. Thereafter, case rates were significantly lower from April 2021 onwards, coinciding with the period of eligibility for NYC adults to be vaccinated against Covid-19. 

<br>  

# Borough by sex, age, and race
```{r, echo = FALSE, message = FALSE, warning = FALSE}
boroughs_by_sex_df = 
  boroughs_by_sex %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bk_case_count:si_death_rate,
    names_to = c("borough", "measure"),
    names_sep = 2
  ) %>% 
  mutate(
    measure = sub("_","",measure),
    borough = recode(borough, "bx" = "Bronx", "bk" = "Brooklyn", "mn" = "Manhattan", "qn" = "Queens", "si" = "Staten Island")
  ) %>% 
  filter(
    measure %in% c("case_count", "hospitalized_count", "death_count"),
    group != "Boroughwide"
    ) 

boroughs_by_age_df =  
 boroughs_by_age %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bk_case_count:si_death_rate,
    names_to = c("borough", "measure"),
    names_sep = 2
  ) %>% 
  mutate(
    measure = sub("_","",measure),
    borough = recode(borough, "bx" = "Bronx", "bk" = "Brooklyn", "mn" = "Manhattan", "qn" = "Queens", "si" = "Staten Island")
  ) %>% 
  filter(
    measure %in% c("case_count", "hospitalized_count", "death_count"),
    group != "Boroughwide"
    ) 

boroughs_by_race_df = 
  boroughs_by_race %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bk_case_count:si_death_rate_adj,
    names_to = c("borough", "measure"),
    names_sep = 2
  ) %>% 
  mutate(
    measure = sub("_","",measure),
    borough = recode(borough, "bx" = "Bronx", "bk" = "Brooklyn", "mn" = "Manhattan", "qn" = "Queens", "si" = "Staten Island")
  ) %>% 
  filter(
    measure %in% c("case_count", "hospitalized_count", "death_count"),
    group != "Boroughwide"
    ) 

boroughs_by_sex_plot =
  boroughs_by_sex_df %>% 
  ggplot(aes(x = group, y = value, fill = measure)) +
    geom_bar(stat = "identity") +
    facet_grid(. ~ borough) +
  labs(
    title = "Plot A: Case, death, hospitalization counts by sex and borough",
    x = "",
    y = "Count"
  )

boroughs_by_age_plot =
  boroughs_by_age_df %>% 
  ggplot(aes(x = group, y = value, fill = measure)) +
    geom_bar(stat = "identity")  +
    facet_grid(. ~ borough) +
  labs(
    title = "Plot B: Case, death, hospitalization counts by age and borough",
    x = "",
    y = "Count"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

boroughs_by_race_plot = 
boroughs_by_race_df %>% 
  ggplot(aes(x = group, y = value, fill = measure)) +
    geom_bar(stat = "identity") +
    facet_grid(. ~ borough) +
  labs(
    title = "Plot C: Case, death, hospitalization counts by race and borough",
    x = "",
    y = "Count"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

boroughs_by_sex_plot
boroughs_by_age_plot
boroughs_by_race_plot
```

The chart in Plot A shows aggregate counts for case, hospitalization and death counts for Covid 19 delineated by sex for all boroughs. It is clear that Queens, Brooklyn and Bronx have the highest aggregate counts in that order of decreasing magnitude. This pattern may explain the clustering of vaccination locations in these boroughs that we explore in subsequent paragraphs. It also noted that the counts for males are consistently higher than for females among all boroughs. 

<br> 

The chart in Plot B shows aggregate counts for case, hospitalization and death counts for Covid 19 delineated by arbitrarily defined age brackets for all boroughs.The greatest concentration of aggregate counts is consistently in the 18-44 and 45-64 year old age bracket which together constitute early and mid adulthood. This is against the general perception that Covid-19 predominantly affects the elderly. This bears significance in public health communication to the public with regards to risk perception.

<br> 

The chart in Plot C shows aggregate counts for case, hospitalization and death counts for Covid 19 delineated by arbitrarily defined race. The chart shows that the greatest burden of the aggregate measures is in the non-white population,  lending further credence to the underlying racial disparities in Covid-19 outcomes and the disproportionate burden of the disease. 

<br> 
    
```{r, echo = FALSE, message = FALSE, warning = FALSE}
partial_vax_plot =
  vax_by_zip_adults_df %>% 
  filter(perc_partially < 30) %>% 
  ggplot(aes(x = boro, y = perc_partially, fill = boro)) +
  geom_boxplot() +
  labs(
    title = "Percent adults partially vaccinated by borough, as of 12/7/2021",
    x = "",
    y = "Percent partially vaccinated"
  )

full_vax_plot =
  vax_by_zip_adults_df %>% 
  filter(perc_fully < 100) %>% 
  ggplot(aes(x = boro, y = perc_fully, fill = boro)) +
  geom_boxplot() +
  labs(
    title = "Percent adults fully vaccinated by borough, as of 12/7/2021",
    x = "",
    y = "Percent fully vaccinated"
  )

partial_vax_plot
full_vax_plot
```

<br> 



These plots show the percent of adults partially and fully vaccinated by borough, as of December 2021. The partially vaccinated plot excluded outliers greater than 30% and the fully vaccinated plot excluded outliers greater than 100%, as they were not representative of the dataset and may have been included erroneously. The median full-vaccination rate is greater than 75% for each borough, with Queens, Manhattan, and Staten Island having the highest full-vaccination rates, respectively. Interestingly, while Staten Island has one of the higher vaccination rates, it is consistently the borough with the highest case rate out of all the boroughs. When we examine the partially and fully vaccinated rates cumulatively, each borough is at least 80% vaccinated. 

<br> 

### We then took a look at our flu data:

On completing exploratory analyses of Covid related data, we then took a look at our flu data:

<br> 

Loading NYC Locations Providing Seasonal Flu Vaccinations/ Emergency Department Visits and Admissions for Influenza-like Illness/ Census Selected Social Characteristics

```{r, message = FALSE, warning = FALSE}

flu_vaxx_loc = read_csv("./data/New_York_City_Locations_Providing_Seasonal_Flu_Vaccinations.csv")

census_social = read_csv("./data/censuszip_selected_social_characterisitcs_new.csv")

```

 
# Data Cleaning 2
Cleaning flu_vaxx_loc:
```{r}

flu_vaxx_byzipcode =
  flu_vaxx_loc %>%
  janitor::clean_names() %>%
  select(walk_in:zip_code, location) %>%
  filter(borough != "YONKERS") %>%
  group_by(zip_code) %>%
  summarize(n_loc = n()) #to get the total number of vax locations by zipcode

flu_vaxx_byboro =
    flu_vaxx_loc %>%
    janitor::clean_names() %>%
    mutate(borough = toupper(borough)) %>%
    filter(borough != "YONKERS") %>%
    group_by(borough) %>%
    summarize(n_loc_boro = n()) #to get the total number of vax locations by zipcode

```

 

## Exploratory Analysis
```{r}
ggplot(flu_vaxx_byboro, aes(x = borough, y = n_loc_boro, fill = borough)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Distribution of Locations Providing
  Seasonal Flu Vaccinations in NYC",
    x = "Boroughs of NYC",
    y = "Number of Locations Providing Seasonal Flu Vaccinations"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
```

The above bar graph shows the number of locations providing the seasonal flu vaccine across all five boroughs of NYC. Based on the graph, we can see that Manhattan has by far the largest number of locations providing the flu vaccine, followed by Brooklyn, Queens, the Bronx, and Staten Island, respectively. 

## Interactive Map of Flu Vaccine Locations Across Five Boroughs in NYC

```{r}
merged_vax = left_join(vax_by_zip_adults_df, zip_lat_lon, by = "zipcode")
```

```{r, message=FALSE}
pal = colorFactor("viridis", NULL)
merged_vax %>%
  leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(~lng, ~lat, radius = 5, color = ~pal(perc_fully), popup = ~perc_fully) %>%
    addLegend("bottomright", pal = pal, values = ~perc_fully,
    title = "Percent fully vaccinated",
    opacity = 1)
```

The interactive map depicts the count of individuals who are fully vaccinated in New York City, categorized by zipcode. Based on the map, we can see that Manhattan and Queens seem to be the areas that have a higher count of individuals who are fully vaccinated. It appears that the Bronx and Brooklyn have a much lower count of individuals who are fully vaccinated. 

```{r, message=FALSE}
pal = colorFactor("viridis", NULL)
merged_vax %>%
  leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(~lng, ~lat, radius = 5, color = ~pal(perc_partially), popup = ~perc_partially) %>%
    addLegend("bottomright", pal = pal, values = ~perc_partially,
    title = "Percent partially vaccinated",
    opacity = 1)
```
The interactive map depicts the count of individuals who are partially vaccinated in New York City, categorized by zipcode. It appears that Queens and Brooklyn (and some parts of the Bronx) have a higher count of individuals who are only partially vaccinated, compared to Manhattan. 
