---
title: "Data Analysis"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---


# Clustering analysis
```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(maptools)
library(spatstat)
library(sf)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .9,
  out.width = "90%"
)
```

## FLU CLUSTERING
```{r, echo = FALSE, message = FALSE, warning = FALSE}
mydata <- read.csv("./data/NYC_Flu_locations.csv")
NYC_Flu_Locations <- ppp(mydata[,13], mydata[,12], c(-74.23442287,-73.68843386), c(40.51789347,40.90856191))
plot(NYC_Flu_Locations)
```

# This is running the K function test
```{r, echo = FALSE, message = FALSE, warning = FALSE}
L <- envelope(NYC_Flu_Locations, Lest, nsim = 10, rank = 2, global = T)
```

# Plot L function with Main Title and x-/y-axis labels
```{r, echo = FALSE, message = FALSE, warning = FALSE}
plot(L, 
     main = "NYC Flu Locations Clustering Results", 
     ylab = "Density Function",
     xlab = "Distance")
```


## COVID CLUSTERING
```{r, echo = FALSE, message = FALSE, warning = FALSE}
mydata2 <- read.csv("./data/COVID_NYC_locations_1.csv") 
NYC_COVID_Locations <- ppp(mydata2[,6], mydata2[,5], c(-74.23538,-73.71330), c(40.51695749,40.90388332))
plot(NYC_COVID_Locations)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
L <- envelope(NYC_COVID_Locations, Lest, nsim = 10, rank = 2, global = T)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
plot(L, 
     main = "NYC COVID Vaccine Locations Clustering Results", 
     ylab = "Density Function",
     xlab = "Distance")
```

The number of locations for Flu vaccinations is a bit more densely populated in lower and mid Manhattan compared to the same area in the number of locations for COVID vaccinations. Comparatively, we see that there are generally more locations for COVID vaccinations in Queens, especially in Far Rockaway, and in Brooklyn. This could be due to the demographic of the population living in these areas, as COVID-19 has disproportionately affected Black communities. 

The K function is used to identify clustering, randomness, or regularity among events in a spatial point pattern. Based on both Ripley K graphs, we see that at all distances, the Lobs(r), which is the observed spatial pattern, is higher than Ltheo(r) , which is the expected random spatial pattern. The Lobs(r) also does not intersect at any point with the expected line. This shows there is statistically significant clustering at smaller distances, displaying a clustering pattern. Due to this clustering pattern, there is not a uniform distribution and we see that there are groups of more COVID and Flu vaccination locations in specific areas while other areas like Staten Island and the parts of Queens are lacking. However, NYCHOH has done a good job on utilizing neighborhood and family owned pharmacies, corporation pharmacies such as Walgreens, as well as  mobile vans (for COVID vaccines), to provide access to the vaccines across NYC. 