---
title: "Project Report"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---
<br></br>

# Introduction: 
<br></br>

### Motivation

As we enter the influenza season this Winter while continuing to fight against the COVID-19 pandemic, we have seen a continued urgency from public health and government officials alike, to encourage New Yorkers to get vaccinated for both influenza and COVID-19. This year in particular, experts are concerned that this year's flu season may be even more severe, due to an unusually mild season last year (Spantideas et al, 2021). Now, with the rise of a new COVID variant, city and public health leaders are pushing for mandatory vaccination requirements and continued COVID-19 precautions to prepare for a potential “Twindemic,” or the simultaneous outbreak of influenza and COVID-19. A Twindemic has the potential to crowd emergency rooms, threaten the public health of our communities, and burden our healthcare system. Therefore, for this project, we seek to conduct an exploratory analysis between COVID-19 and influenza in New York City.  Specifically, we will examine the relationship between COVID-19 and influenza uptake by zip code and sociodemographic factors in New York City through data visualizations and clustering analyses. 
 
### Related work: 

New York City was the epicenter of  COVID-19 during the early stages of the COVID-19 pandemic, with 12,589.2 cases and 404.5 deaths per 100,000 residents recorded in September of 2021 (Moreland et al., 2021). Recently, alongside entering the winter season and the emergence of yet another COVID-19 variant, news outlets and academic sources alike have warned of a threat of a Twindemic. The urgency to get vaccinated for not only COVID-19, but also influenza, is prevalent now more than ever among city officials and public health professionals to curb the spread of two infectious diseases. The COVID-19 pandemic shed light on stark racial/ethnic health disparities related to COVID-19 illness and mortality. Furthermore, vaccine hesitancy has exacerbated these disparities, with more people of color disproportionately affected. We aim for the results of our analyses of the COVID-19 and influenza trends in New York City to inform public health programs and policies to improve vaccination efforts to vulnerable communities. 
 
### Initial questions
A few main questions we had for our exploratory analysis on COVID-19 and influenza in New York City included: 
<li>What is the relationship between COVID and Flu Vaccination Uptake by Zipcode and Sociodemographic Factors in NYC?
<li>What is the difference in COVID and Flu vaccine uptake in NYC, according to  associated sociodemographic factors?
<li>What are the patterns observed for COVID and Flu vaccine uptake in New York by zip code?
<li>Difference in COVID and Flu vaccine uptake in NYC and associated sociodemographic factors.
 
 
# Data
<br></br>

### NYC Health COVID-19 

The NYC Department of Health and Mental Hygiene (NYC DOH) classifies the start of the COVID-19 outbreak in NYC by its first lab-confirmed case on February 29, 2020. Cases of COVID-19 included in their datasets are considered “confirmed” cases, or cases with a positive result from a molecular test. NYC DOH data on COVID-19 vaccines come from the Citywide Immunization Registry (CIR), which tracks immunizations for all NYC residents. A person is considered fully vaccinated if they have received both doses of the Moderna or Pfizer vaccine series, or received one dose of the Johnson & Johnson/Janssen vaccine series. There is currently no data on COVID-19 boosters. The NYC DOH COVID-19 data can be found here, and the COVID-19 vaccination data can be found here. 

### New York City Locations Providing Seasonal Flu Vaccine
<br></br>
https://data.cityofnewyork.us/Health/New-York-City-Locations-Providing-Seasonal-Flu-Vac/w9ei-idxz

### Census Data

ZIP Code Data
We also identified a datase that provided the corresponding latitude and longitude for each ZIP code, which we filtered to only include the ZIP codes relevant to our analyses of NYC.

https://gist.github.com/erichurst/7882666

 
Restricting on adults since vaccines weren’t available to children until recently. Therefore, we want to examine the COVID-19 outcomes among adults in this analysis. 
Exploratory analysis: Visualizations, summaries, and exploratory statistical analyses. Justify the steps you took, and show any major changes to your ideas.
 
# Exploratory Analysis 
During our exploratory analysis, we first examined COVID-19 and influenza case and vaccine trends separately to better understand the data we were working with. 
COVID-19 

```{r setup, include = FALSE}
library(tidyverse)
library(dplyr)
library(readxl)
library(haven)
library(ggplot2)
library(patchwork)
library(leaflet)


knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

### Data Cleaning

Loading the COVID data from NYC Health.
```{r, message = FALSE}
caserate_by_zip = read_csv("./data/caserate-by-modzcta.csv")
vax_by_boro_age = read_csv("./data/coverage-by-boro-age.csv")
vax_by_boro_demo = read_csv("./data/coverage-by-boro-demo.csv")
data_by_day = read_csv("./data/data-by-day.csv")
doses_by_day = read_csv("./data/doses-by-day.csv")
hosprate_by_zip = read_csv("./data/hosprate-by-modzcta.csv")
testrate_by_zip = read_csv("./data/testrate-by-modzcta.csv")
vax_by_zip_adults = read_csv("./data/coverage-by-modzcta-adults.csv")

boroughs_by_sex = read_csv("./data/boroughs-by-sex.csv")
boroughs_by_age = read_csv("./data/boroughs-by-age.csv")
boroughs_by_race = read_csv("./data/boroughs-by-race.csv")


zip_lat_lon = 
  read.table("./data/us_zipcode_lat_lon.txt", sep = ",", header = TRUE) %>% 
  janitor::clean_names() %>% 
  rename(zipcode = zip) %>% 
  filter(zipcode %in% c(10001:11697))
```
From the datasets above, we will likely compare the following pairs: caserate/hosprate/testrate_by_zip; data/doses_by_day; vax_by_boro_age/demo



Cleaning caserate_by_zip:
```{r}
caserate_by_zipcode = 
  caserate_by_zip %>% 
  janitor::clean_names() %>% 
  select(week_ending, caserate_10001:caserate_10280) %>% 
  pivot_longer(
    caserate_10001:caserate_10280,
    names_to = "zipcode", 
    names_prefix = "caserate_",
    values_to = "case_rate"
  ) %>% 
  mutate(
    week_ending = as.Date(week_ending, "%m/%d/%Y")) 

caserate_by_boro = 
  caserate_by_zip %>% 
  janitor::clean_names() %>% 
  select(week_ending, caserate_city:caserate_si) %>% 
  pivot_longer(
    caserate_city:caserate_si,
    names_to = "boro", 
    names_prefix = "caserate_",
    values_to = "case_rate"
  ) %>% 
  mutate(
    week_ending = as.Date(week_ending, "%m/%d/%Y"),
    boro = recode(boro, "bx" = "Bronx", "bk" = "Brooklyn", "mn" = "Manhattan", "qn" = "Queens", "si" = "Staten Island")) 
```


Cleaning testrate_by_zip:
```{r}
testrate_by_zipcode = 
  testrate_by_zip %>% 
  janitor::clean_names() %>% 
  select(week_ending, testrate_10001:testrate_10280) %>% 
  pivot_longer(
    testrate_10001:testrate_10280,
    names_to = "zipcode", 
    names_prefix = "testrate_",
    values_to = "test_rate"
  ) %>% 
  mutate(
    week_ending = as.Date(week_ending, "%m/%d/%Y")) 

testrate_by_boro = 
  testrate_by_zip %>% 
  janitor::clean_names() %>% 
  select(week_ending, testrate_city:testrate_si) %>% 
  pivot_longer(
    testrate_city:testrate_si,
    names_to = "boro", 
    names_prefix = "testrate_",
    values_to = "test_rate"
  ) %>% 
  mutate(
    week_ending = as.Date(week_ending, "%m/%d/%Y"),
    boro = recode(boro, "bx" = "Bronx", "bk" = "Brooklyn", "mn" = "Manhattan", "qn" = "Queens", "si" = "Staten Island")) 
```


Cleaning hosprate_by_zip:
```{r}
hosprate_by_zipcode = 
  hosprate_by_zip %>% 
  janitor::clean_names() %>% 
  select(date, hosprate_10001:hosprate_10280) %>% 
  pivot_longer(
    hosprate_10001:hosprate_10280,
    names_to = "zipcode", 
    names_prefix = "hosprate_",
    values_to = "hosp_rate"
  )

hosprate_by_boro = 
  hosprate_by_zip %>% 
  janitor::clean_names() %>% 
  select(date, hosprate_bronx:hosprate_citywide) %>% 
  pivot_longer(
    hosprate_bronx:hosprate_citywide,
    names_to = "boro", 
    names_prefix = "hosprate_",
    values_to = "hosp_rate"
  ) %>% 
  mutate(
    boro = recode(boro, "bronx" = "Bronx", "brooklyn" = "Brooklyn", "manhattan" = "Manhattan", "queens" = "Queens", "staten_island" = "Staten Island")
  )
```


Cleaning vax_by_boro_age/demo:
```{r}
vax_by_boro_age_df = 
  vax_by_boro_age %>% 
  janitor::clean_names() %>% 
  filter(age_group %in% c("'18-24", "'25-34", "'35-44", "'45-54", "'55-64", "'65-74", "'75-84", "'85+") )

vax_by_boro_demo_df = 
  vax_by_boro_demo %>% 
  janitor::clean_names() %>% 
  filter(age_group %in% c("'18-44", "'45-64", "'65+", "All ages") )
```

Cleaning data/doses_by_day:
```{r}
doses_by_day_df = 
  doses_by_day %>% 
  janitor::clean_names()

data_by_day_df = 
  data_by_day %>% 
  janitor::clean_names() %>% 
  mutate(
    date_of_interest = as.Date(date_of_interest, "%m/%d/%Y")
  )
```

Cleaning vax_by_zip_adults:
```{r}
vax_by_zip_adults_df = 
  vax_by_zip_adults %>% 
  janitor::clean_names() %>% 
  rename(boro = borough, zipcode = modzcta)
```

## Exploratory Analysis: COVID-19 Rates & Vaccinations
```{r, echo = FALSE, message = FALSE, warning = FALSE}
caserate_by_boro_plot = 
  caserate_by_boro %>% 
  filter(boro != "city") %>% 
  ggplot(aes(x = week_ending, y = case_rate, color = boro)) +
  geom_line() +
  labs(
    title = "Weekly Case Rate by Borough",
    x = "",
    y = "Case Rate"
  )

caserate_by_boro_plot
```

COVID-19 
<br></br>

This plot shows the weekly case rate by NYC borough, beginning the week starting on August 9, 2020. The plot shows a similar trend in weekly rates across all 5 boroughs. Cases peaked for all boroughs from January through February of 2021. Boroughs experienced a smaller peak in cases in April 2021. Thereafter, case rates were significantly lower from April 2021 onwards, coinciding with the period of eligibility for NYC adults to be vaccinated against Covid-19. 

<br>  

# Borough by sex, age, and race
```{r, echo = FALSE, message = FALSE, warning = FALSE}
boroughs_by_sex_df = 
  boroughs_by_sex %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bk_case_count:si_death_rate,
    names_to = c("borough", "measure"),
    names_sep = 2
  ) %>% 
  mutate(
    measure = sub("_","",measure),
    borough = recode(borough, "bx" = "Bronx", "bk" = "Brooklyn", "mn" = "Manhattan", "qn" = "Queens", "si" = "Staten Island")
  ) %>% 
  filter(
    measure %in% c("case_count", "hospitalized_count", "death_count"),
    group != "Boroughwide"
    ) 

boroughs_by_age_df =  
 boroughs_by_age %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bk_case_count:si_death_rate,
    names_to = c("borough", "measure"),
    names_sep = 2
  ) %>% 
  mutate(
    measure = sub("_","",measure),
    borough = recode(borough, "bx" = "Bronx", "bk" = "Brooklyn", "mn" = "Manhattan", "qn" = "Queens", "si" = "Staten Island")
  ) %>% 
  filter(
    measure %in% c("case_count", "hospitalized_count", "death_count"),
    group != "Boroughwide"
    ) 

boroughs_by_race_df = 
  boroughs_by_race %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bk_case_count:si_death_rate_adj,
    names_to = c("borough", "measure"),
    names_sep = 2
  ) %>% 
  mutate(
    measure = sub("_","",measure),
    borough = recode(borough, "bx" = "Bronx", "bk" = "Brooklyn", "mn" = "Manhattan", "qn" = "Queens", "si" = "Staten Island")
  ) %>% 
  filter(
    measure %in% c("case_count", "hospitalized_count", "death_count"),
    group != "Boroughwide"
    ) 

boroughs_by_sex_plot =
  boroughs_by_sex_df %>% 
  ggplot(aes(x = group, y = value, fill = measure)) +
    geom_bar(stat = "identity") +
    facet_grid(. ~ borough) +
  labs(
    title = "Plot A: Case, death, hospitalization counts by sex and borough",
    x = "",
    y = "Count"
  )

boroughs_by_age_plot =
  boroughs_by_age_df %>% 
  ggplot(aes(x = group, y = value, fill = measure)) +
    geom_bar(stat = "identity")  +
    facet_grid(. ~ borough) +
  labs(
    title = "Plot B: Case, death, hospitalization counts by age and borough",
    x = "",
    y = "Count"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

boroughs_by_race_plot = 
boroughs_by_race_df %>% 
  ggplot(aes(x = group, y = value, fill = measure)) +
    geom_bar(stat = "identity") +
    facet_grid(. ~ borough) +
  labs(
    title = "Plot C: Case, death, hospitalization counts by race and borough",
    x = "",
    y = "Count"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

boroughs_by_sex_plot
boroughs_by_age_plot
boroughs_by_race_plot
```

The chart in Plot A shows aggregate counts for case, hospitalization and death counts for Covid 19 delineated by sex for all boroughs. It is clear that Queens, Brooklyn and Bronx have the highest aggregate counts in that order of decreasing magnitude. This pattern may explain the clustering of vaccination locations in these boroughs that we explore in subsequent paragraphs. It also noted that the counts for males are consistently higher than for females among all boroughs. 

<br> 

The chart in Plot B shows aggregate counts for case, hospitalization and death counts for Covid 19 delineated by arbitrarily defined age brackets for all boroughs. The greatest concentration of aggregate counts is consistently in the 18-44 and 45-64 year old age bracket which together constitute early and mid adulthood. This is against the general perception that Covid-19 predominantly affects the elderly. This bears significance in public health communication to the public with regards to risk perception.

<br> 

The chart in Plot C shows aggregate counts for case, hospitalization and death counts for Covid 19 delineated by arbitrarily defined race. The chart shows that the greatest burden of the aggregate measures is in the non-white population,  lending further credence to the underlying racial disparities in Covid-19 outcomes and the disproportionate burden of the disease. 


<br> 

```{r, echo = FALSE, message = FALSE, warning = FALSE}
partial_vax_plot =
  vax_by_zip_adults_df %>% 
  filter(perc_partially < 30) %>% 
  ggplot(aes(x = boro, y = perc_partially, fill = boro)) +
  geom_boxplot() +
  labs(
    title = "Percent adults partially vaccinated by borough, as of 12/7/2021",
    x = "",
    y = "Percent partially vaccinated"
  )

full_vax_plot =
  vax_by_zip_adults_df %>% 
  filter(perc_fully < 100) %>% 
  ggplot(aes(x = boro, y = perc_fully, fill = boro)) +
  geom_boxplot() +
  labs(
    title = "Percent adults fully vaccinated by borough, as of 12/7/2021",
    x = "",
    y = "Percent fully vaccinated"
  )

partial_vax_plot
full_vax_plot
```

These plots show the percent of adults partially and fully vaccinated by borough, as of December 2021. The partially vaccinated plot excluded outliers greater than 30% and the fully vaccinated plot excluded outliers greater than 100%, as they were not representative of the dataset and may have been included erroneously. The median full-vaccination rate is greater than 75% for each borough, with Queens, Manhattan, and Staten Island having the highest full-vaccination rates, respectively. Interestingly, while Staten Island has one of the higher vaccination rates, it is consistently the borough with the highest case rate out of all the boroughs. When we examine the partially and fully vaccinated rates cumulatively, each borough is at least 80% vaccinated. 


## Interactive Map of Full Vaccination by Zipcode in NYC 


```{r}
merged_vax = left_join(vax_by_zip_adults_df, zip_lat_lon, by = "zipcode")
```

```{r, message=FALSE}
pal = colorFactor("viridis", NULL)
merged_vax %>%
  leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(~lng, ~lat, radius = 5, color = ~pal(perc_fully), popup = ~perc_fully) %>%
    addLegend("bottomright", pal = pal, values = ~perc_fully,
    title = "Percent fully vaccinated",
    opacity = 1)
```
The interactive map depicts the count of individuals who are fully vaccinated in New York City, categorized by zipcode. Based on the map, we can see that Manhattan and Queens seem to be the areas that have a higher count of individuals who are fully vaccinated. It appears that the Bronx and Brooklyn have a much lower count of individuals who are fully vaccinated. 


## Interactive Map of Partial Vaccination by Zipcode in NYC

```{r, message=FALSE}
pal = colorFactor("viridis", NULL)
merged_vax %>%
  leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(~lng, ~lat, radius = 5, color = ~pal(perc_partially), popup = ~perc_partially) %>%
    addLegend("bottomright", pal = pal, values = ~perc_partially,
    title = "Percent partially vaccinated",
    opacity = 1)
```
The interactive map depicts the count of individuals who are partially vaccinated in New York City, categorized by zipcode. It appears that Queens and Brooklyn (and some parts of the Bronx) have a higher count of individuals who are only partially vaccinated, compared to Manhattan. 
<br> 

# We then took a look at our flu data:
Loading NYC Locations Providing Seasonal Flu Vaccinations/ Emergency Department Visits and Admissions for Influenza-like Illness/ Census Selected Social Characteristics

```{r, message = FALSE, warning = FALSE}

flu_vaxx_loc = read_csv("./data/New_York_City_Locations_Providing_Seasonal_Flu_Vaccinations.csv")

census_social = read_csv("./data/censuszip_selected_social_characterisitcs_new.csv")

```

 
## Data Cleaning 2
Cleaning flu_vaxx_loc:
```{r}

flu_vaxx_byzipcode =
  flu_vaxx_loc %>%
  janitor::clean_names() %>%
  select(walk_in:zip_code, location) %>%
  filter(borough != "YONKERS") %>%
  group_by(zip_code) %>%
  summarize(n_loc = n()) #to get the total number of vax locations by zipcode

flu_vaxx_byboro =
    flu_vaxx_loc %>%
    janitor::clean_names() %>%
    mutate(borough = toupper(borough)) %>%
    filter(borough != "YONKERS") %>%
    group_by(borough) %>%
    summarize(n_loc_boro = n()) #to get the total number of vax locations by zipcode

```

 

## Exploratory Analysis

```{r}
ggplot(flu_vaxx_byboro, aes(x = borough, y = n_loc_boro, fill = borough)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Distribution of Locations Providing
  Seasonal Flu Vaccinations in NYC",
    x = "Boroughs of NYC",
    y = "Number of Locations Providing Seasonal Flu Vaccinations"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
```
The above bar graph shows the number of locations providing the seasonal flu vaccine across all five boroughs of NYC. Based on the graph, we can see that Manhattan has by far the largest number of locations providing the flu vaccine, followed by Brooklyn, Queens, the Bronx, and Staten Island, respectively. 



# Additional analysis
## Clustering analysis

```{r}
library(maptools)
library(spatstat)
library(sf)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .9,
  out.width = "90%"
)
```

## FLU CLUSTERING
```{r}
mydata <- read.csv("./data/NYC_Flu_locations.csv")
NYC_Flu_Locations <- ppp(mydata[,13], mydata[,12], c(-74.23442287,-73.68843386), c(40.51789347,40.90856191))
plot(NYC_Flu_Locations)
```

## K function test
```{r}
L <- envelope(NYC_Flu_Locations, Lest, nsim = 10, rank = 2, global = T)
```

## Plot L function with Main Title and x-/y-axis labels
```{r}
plot(L, 
     main = "NYC Flu Locations Clustering Results", 
     ylab = "Density Function",
     xlab = "Distance")
```


## COVID CLUSTERING
```{r}
mydata2 <- read.csv("./data/COVID_NYC_locations_1.csv") 
NYC_COVID_Locations <- ppp(mydata2[,6], mydata2[,5], c(-74.23538,-73.71330), c(40.51695749,40.90388332))
plot(NYC_COVID_Locations)
```

```{r}
L <- envelope(NYC_COVID_Locations, Lest, nsim = 10, rank = 2, global = T)
```

```{r}
plot(L, 
     main = "NYC COVID Vaccine Locations Clustering Results", 
     ylab = "Density Function",
     xlab = "Distance")
```

The number of locations for Flu vaccinations is a bit more densely populated in lower and mid Manhattan compared to the same area in the number of locations for COVID vaccinations. Comparatively, we see that there are generally more locations for COVID vaccinations in Queens, especially in Far Rockaway, and in Brooklyn. This could be due to the demographic of the population living in these areas, as COVID-19 has disproportionately affected Black communities. 

The K function is used to identify clustering, randomness, or regularity among events in a spatial point pattern. Based on both Ripley K graphs, we see that at all distances, the Lobs(r), which is the observed spatial pattern, is higher than Ltheo(r) , which is the expected random spatial pattern. The Lobs(r) also does not intersect at any point with the expected line. This shows there is statistically significant clustering at smaller distances, displaying a clustering pattern. Due to this clustering pattern, there is not a uniform distribution and we see that there are groups of more COVID and Flu vaccination locations in specific areas while other areas like Staten Island and the parts of Queens are lacking. However, NYCHOH has done a good job on utilizing neighborhood and family owned pharmacies, corporation pharmacies such as Walgreens, as well as  mobile vans (for COVID vaccines), to provide access to the vaccines across NYC. 
 
# Discussion
The exploratory and clustering analyses of COVID-19 rates and vaccinations in New York City reveal several findings that are important to note. First of all, as we can see from the exploratory analyses, the COVID-19 case, death, and hospitalization counts are higher in Queens, Brooklyn, and the Bronx, with higher counts among men (compared to women), individuals in the 18-44 and 45-64 age groups, and among individuals of color in all five boroughs except for Staten Island. Our exploratory findings highlight and confirm the disparity in COVID-19 disease burden that we have seen across New York City. 

Furthermore, in terms of vaccination, our analysis reveals that the median full-vaccination rate is greater than 75% in all boroughs, with Queens, Manhattan, and Staten Island, with the highest vaccination rates, respectively. Interestingly, while Staten Island had one the highest vaccination rates, it also has the highest COVID-19 case rates compared to all other boroughs. In addition, we found that vaccination counts were highest among whites in three out of five boroughs. This is notably important to address in public health risk communication strategies to attempt to increase vaccine uptake in more vulnerable communities.

Our analysis of the flu vaccinations indicate that Manhattan, Brooklyn, and Queens, respectively, are the boroughs with the highest distribution of seasonal flu vaccination locations. The interactive map of flu vaccine locations across all five boroughs in New York City further reveal the disparities in access to flu vaccines, with Manhattan having by far the largest number of flu vaccination locations, followed by Brooklyn, Queens, Bronx, and Staten Island, respectively. 

Lastly, we conducted a cluster analysis for both flu and COVID-19 vaccine locations to compare the two. The clustering analysis revealed that flu vaccine locations were more prevalent in lower and mid-Manhattan compared the COVID-19 vaccine locations in the same area. Furthermore, we can see that there appear to be more COVID-19 vaccination sites in Queens (especially Far Rockaway) and Brooklyn. This could be due to public health efforts that aimed to reduce disease burden in those neighborhoods/communities since these areas were more severely impacted by COVID-19 compared to other communities. 

Using the K function to determine the statistical significance of our clustering findings, since the Lobs(r) (the observed spatial pattern) does not intersect with the expected line at any point, we have evidence of statistically significant clustering at smaller distances, which displays a clustering pattern. This clustering pattern shows us there is not a uniform distribution of COVID-19 and flu vaccination sites, with certain areas in New York City with a higher concentration of vaccination sites compared to others. Our data confirms that the NYC Health + Hospitals (NYCHOH) has utilized various types of healthcare distribution centers (such as family pharmacies, mobile vans, corporation pharmacies, etc) to provide wide scale access to vaccines in New York City. 

Our findings indicate that there is much work to be done when it comes to flu and vaccine equity in New York City. The disparities in COVID-19 case, death, and hospitalization counts in all five boroughs reveal stark disparities in our city that we, as public health professionals, must work to address. Initiatives to address these health disparities in our city include allocating more resources towards flu and COVID-19 vaccination sites in areas most disproportionately affected by COVID-19 and flu outbreaks. Furthermore, it is imperative to also launch community engagement and educational initiatives to combat misinformation and vaccine hesitancy  in order to ensure that we are promoting public health and health equity for all. 

The group members of this project represent diverse fields of public health, including epidemiology, biostatistics, public health research methods, and health policy. We all hope to advance health care equity by bridging data science into practice, policy, and health and human rights for all. 


# Limitation
Flu and covid vaccine data did not overlap, explain
